---
title: "NCD_HIV SystematicReview_Code"
author: "JMG"
date: "2022-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Loading libraries

```{r}
library(readxl)
library(ggplot2)
library(ggpubr)
library(meta)
library(metasens)
library(metafor)
library(metaforest)
library(dplyr)
```

## Loading data sets

```{r}
## Dataset with obesity articles
Obes <- read_excel("D:/AXA.Researh/ScR/AXA_NCD/NCD_HIV/Datasets/NCD_HIV_rawdata.xlsx", 
                                              sheet = "DA1")
View(Obes)

## Dataset with HPT articles
HPT <- read_excel("D:/AXA.Researh/ScR/AXA_NCD/NCD_HIV/Datasets/NCD_HIV_rawdata.xlsx", 
                                              sheet = "DA2")
View(HPT)

## Dataset with Diabetes articles
diabetes <- read_excel("D:/AXA.Researh/ScR/AXA_NCD/NCD_HIV/Datasets/NCD_HIV_rawdata.xlsx", 
                                              sheet = "DA3")
View(diabetes)
```

## Combine all the 3 datasets

```{r}
NCD.dat <- rbind(Obes, HPT, diabetes)
View(NCD.dat)

str(NCD.dat) #explore structure of the combined dataset

NCD.dat[,c(1,2,4:8,11:14)] <- lapply(NCD.dat[,c(1,2,4:8,11:14)] , factor) #coerce some variables

summary(NCD.dat) ##summary statistics


```
## Calculate the prevalence and 95% CI
```{r}
## calculate the prevalence
NCD.dat <- NCD.dat%>%
  mutate(Prevalence = round((N/size)*100,1))

## Plot prevalence against Year of data collection for each NCD

plot1<-ggplot(NCD.dat, aes(x=Collection, y=Prevalence)) +
  geom_point() +
  theme_classic()+
  scale_x_continuous(breaks = scales::pretty_breaks(n=8), expand = c(0.1,0))+
  facet_wrap(~NCD, ncol = 2)+
  labs(x="Year of Data collection",y="Prevalence (%)")+
  theme(
    axis.title.x = element_text(size = 9,face = "bold"),
    axis.text.y = element_text(size = 9,face = "bold", color = "black"),
    axis.text.x = element_text(size = 9, face = "bold",color = "black"),
    axis.title.y = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9,face = "bold"),
    strip.text = element_text(size=9, face = "bold"))+
  geom_smooth(method = lm)+
  stat_cor(method = "spearman"); plot1

## save the plot as .tiff flie  
ggsave(plot1, filename = "plot1.tiff", width = 10, height = 6,dpi = 600)


## 95% CI for prevalence
#### margin of error
x<- NCD.dat$Prevalence_2*(1- NCD.dat$Prevalence_2)/NCD.dat$size
y <- sqrt(x)
z <- qnorm(0.975)
marginErr <- y*z
NCD.dat$marginErr <- marginErr
NCD.dat$SE <- round (y*100,3)


## 95% CI lower interval
NCD.dat$Lower_CI <- round((NCD.dat$Prevalence_2 - NCD.dat$marginErr)*100,2)

## 95% CI Upper interval
NCD.dat$Upper_CI <- round((NCD.dat$Prevalence_2 + NCD.dat$marginErr)*100,2)

NCD.dat$index <- seq(1:116)
```

## Plot: Prevalence of each NCD by Author and study setting
```{r}
#### obesity prevalence
obes_plot <-NCD.dat%>%
  filter(NCD=="obesity")%>%
  ggplot(data=., aes(y=index, x=Prevalence, xmin=Lower_CI, xmax=Upper_CI, fill=Setting,
                     color=Setting)) +
  geom_point() +
  scale_fill_manual(values = c("#C3D7A4", "#52854C", "#4E84C4", "#293352"))+
  geom_errorbarh(height=.1) +
  scale_y_continuous(name = "", breaks=1:nrow(NCD.dat), labels=NCD.dat$Author, expand = c(0,0.3))+
  scale_x_continuous(breaks = scales::pretty_breaks(n=8), expand = c(0.1,0))+
  labs(x='Prevalence of obesity', y = 'Study') +
  geom_vline(xintercept=0, color='black', linetype='dashed', alpha=.5) +
  theme_classic()
obes_plot

## save as .tiff file
ggsave(obes_plot, filename = "obes_plot.tiff", width = 10, height = 6,dpi = 600)


#### HPT prevalence
HPT_plot <-NCD.dat%>%
  filter(NCD=="HPT")%>%
  ggplot(data=., aes(y=index, x=Prevalence, xmin=Lower_CI, xmax=Upper_CI, fill=Setting,
                     color=Setting)) +
  geom_point() +
  scale_fill_manual(values = c("#C3D7A4", "#52854C", "#4E84C4", "#293352"))+
  geom_errorbarh(height=.1) +
  scale_y_continuous(name = "", breaks=1:nrow(NCD.dat), labels=NCD.dat$Author, expand = c(0,0.3))+
  scale_x_continuous(breaks = scales::pretty_breaks(n=8), expand = c(0.1,0))+
  labs(x='Prevalence of HPT', y = 'Study') +
  geom_vline(xintercept=0, color='black', linetype='dashed', alpha=.5) +
  theme_classic(); HPT_plot

## save as .tiff file
ggsave(HPT_plot, filename = "HPT_plot.tiff", width = 10, height = 6,dpi = 600)


#### diabetes prevalence
DM_plot <-NCD.dat%>%
  filter(NCD=="diabetes")%>%
  ggplot(data=., aes(y=index, x=Prevalence, xmin=Lower_CI, xmax=Upper_CI, fill=Setting,
                     color=Setting)) +
  geom_point() +
  scale_fill_manual(values = c("#C3D7A4", "#52854C", "#4E84C4", "#293352"))+
  geom_errorbarh(height=.1) +
  scale_y_continuous(name = "", breaks=1:nrow(NCD.dat), labels=NCD.dat$Author, expand = c(0,0.3))+
  scale_x_continuous(breaks = scales::pretty_breaks(n=8), expand = c(0.1,0))+
  labs(x='Prevalence of diabetes', y = 'Study') +
  geom_vline(xintercept=0, color='black', linetype='dashed', alpha=.5) +
  theme_classic()
DM_plot

ggsave(DM_plot, filename = "DM_plot.tiff", width = 10, height = 6,dpi = 600)

```



Summarize and plot number of studies by study setting for each NCD
```{r}
StudiesPerSetting <- NCD.dat%>%
  group_by(NCD,Setting)%>%
  summarise(N=n())
StudiesPerSetting  


plot3 <- ggplot(aes(x=Setting, y=N, fill=NCD), data = StudiesPerSetting)+
  geom_col(position = "dodge", width = 0.5, )+
  theme_classic()+
  scale_fill_manual(values = c("red","blue","green"))+
 labs(x="Study Setting",y="Number of studies")+
 scale_y_continuous(breaks = scales:: pretty_breaks(n=8), expand = c(0,0))+
  theme(
    axis.title.x = element_text(size = 9,face = "bold"),
    axis.text.y = element_text(size = 9,face = "bold", color = "black"),
    axis.text.x = element_text(size = 9, face = "bold",color = "black"),
    axis.title.y = element_text(size = 9, face = "bold"),
    legend.text = element_text(size = 9),
    legend.title = element_text(size = 9,face = "bold")); plot3

## save the plot as .tiff file
ggsave(plot3, filename = "plot3.tiff", width = 10, height = 6,dpi = 600)
```

## Categorize years of data collection: <=2012, and >2012
```{r}
summary(NCD.dat$Collection)
NCD.dat$Study_Period <- ifelse(NCD.dat$Collection<=2012,0,1)
table(NCD.dat$Study_Period)
NCD.dat$Study_Period<- factor(NCD.dat$Study_Period, levels = c(0,1),
                                labels = c("2000-2012","2013-2022"))
```

## Customise the levels in pregnant variable as "NA-(men)","Not pregnant","pregnant/PP"
```{r}
#####pregnancy
levels(NCD.dat$pregnant)
NCD.dat$pregnant <- as.character(NCD.dat$pregnant)
NCD.dat$pregnant[is.na(NCD.dat$pregnant)]<- 0
NCD.dat$pregnant <- as.factor(NCD.dat$pregnant)
NCD.dat$pregnant <- factor(NCD.dat$pregnant, 
                           levels = c("0","no","yes"),
                           labels = c("NA-(men)","Not pregnant","pregnant/PP"))
```

# META-ANALYSIS & FOREST PLOTS
## DIABETES
```{r}
## Subset out articles on diabetes
diabetes.1 <- NCD.dat%>%filter(NCD=="diabetes")

## Summarize the data by sub-groups/ variables of interest
diabetes.1 <- diabetes.1%>%
  group_by(Author, Setting, Province,Study_Period, `Diagnostic criteria`,
           ROB, ART_status)%>%
  summarise(N=sum(N), size=sum(size))
View(diabetes.1) # view the dataset


#### remove studies that did not report ART status unreported
diabetes.ART<-diabetes.1%>%filter(ART_status!="unreported")

##### meta analysis for diabetes: when studies are subgrouped by ART status (initiated Vs not initiated)
pool.DM.ART <- metaprop(event = diabetes.ART$N,n=diabetes.ART$size,
                         fixed = F, random = T, method = "Inverse", method.tau = "DL",
                         sm = "PLOGIT", hakn = T, studlab = diabetes.ART$Author,
                         overall = T, overall.hetstat = T,
                         method.bias = "Egger",backtransf = T,
                         text.random = "RE Prevalence",
                         subgroup = diabetes.ART$ART_status,
                         test.subgroup = T, pscale = 100, )
summary(pool.DM.ART)


###forest plot for diabetes subgroup ART status
ART_forest_DM <- forest.meta(pool.DM.ART,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,40), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

## save the forest plot
png(file = "ART_forest-DM.png", width = 2250, height = 2600, res = 300)
ART_forest_DM
dev.off()


##### meta analysis for diabetes: studies subgrouped by diagnostic criteria
pool.DM.diag <- metaprop(event = diabetes.1$N,n=diabetes.1$size,
                        fixed = F, random = T, method = "Inverse", method.tau = "DL",
                        sm = "PLOGIT", hakn = T, studlab = diabetes.1$Author,
                        overall = T, overall.hetstat = T,
                        method.bias = "Egger",backtransf = T,
                        text.random = "RE Prevalence",
                        subgroup = diabetes.1$`Diagnostic criteria`,
                        test.subgroup = T, pscale = 100, )
summary(pool.DM.diag)

###forest plot for diabetes subgroup diagnostics

Diagnostics_forest_DM <- forest.meta(pool.DM.diag,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,100), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

## save the forest plot
png(file = "Diagnositcs_forest-DM.png", width = 2250, height = 2600, res = 300)
Diagnostics_forest_DM
dev.off()


##### meta analysis for diabetes: studies subgrouped by setting: urban vs rural
diabetes.1.setting <- diabetes.1%>%filter(Setting!="Urban_Rural")
pool.DM.3 <- metaprop(event = diabetes.1.setting$N,n=diabetes.1.setting$size,
                     fixed = F, random = T, method = "Inverse", method.tau = "DL",
                     sm = "PLOGIT", hakn = T, studlab = diabetes.1.setting$Author,
                     overall = T, overall.hetstat = T,
                     method.bias = "Egger",backtransf = T,
                     text.random = "RE Prevalence",
                     subgroup = diabetes.1.setting$Setting,
                    test.subgroup = T, pscale = 100, )
summary(pool.DM.3)



###forest plot for diabetes subgroup setting

forest_setting_DM <- forest.meta(pool.DM.3,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,80), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

## save forest plot
png(file = "forest-DM_2.png", width = 2250, height = 2350, res = 300)
forest_setting_DM
dev.off()


##### meta analysis for diabetes: studies subgrouped by STUDY PERIOD: <=2012 vs >2012
pool.DM.4 <- metaprop(event = diabetes.1$N,n=diabetes.1$size,
                      fixed = F, random = T, method = "Inverse", method.tau = "DL",
                      sm = "PLOGIT", hakn = T, studlab = diabetes.1$Author,
                      overall = T, overall.hetstat = T,
                      method.bias = "Egger",backtransf = T,
                      text.random = "RE Prevalence",
                      subgroup = diabetes.1$Study_Period,
                      test.subgroup = T, pscale = 100, )
summary(pool.DM.4)

###forest plot for diabetes subgroup study period

forest_Studyperiod_DM <- forest.meta(pool.DM.4,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(16,"mm"),
            lwd = 2,colgap.left = unit(1,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(0,45), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

## save forest plot
png(file = "forest-DM_period.png", width = 2250, height = 2350, res = 300)
forest_Studyperiod_DM
dev.off()

#### MA for diabetes: studies subgrouped by sex
DM_sex<- NCD.dat%>%filter(NCD=="diabetes", !is.na(Sex))
View(DM_sex)

pool.DM.sex <- metaprop(event = DM_sex$N,n=DM_sex$size,
                      fixed = F, random = T, method = "Inverse", method.tau = "DL",
                      sm = "PLOGIT", hakn = T, studlab = DM_sex$Author,
                      overall = T, overall.hetstat = T,
                      method.bias = "Egger",backtransf = T,
                      text.random = "RE Prevalence",
                      subgroup = DM_sex$Sex,
                      test.subgroup = T, pscale = 100, )
summary(pool.DM.sex)

## forest plot
forest_sex_DM <- forest.meta(pool.DM.sex,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(16,"mm"),
            lwd = 2,colgap.left = unit(1,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(0,45), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

## save forest plot
png(file = "forest-DM_sex.png", width = 2800, height = 2700, res = 300)
forest_sex_DM
dev.off()


```

## HYPERTENTION
```{r}
### subset out articles on hypertension

HPT.1 <- NCD.dat%>%filter(NCD=="HPT")

#### remove studies on pregnant/postpartum participants
HPT.1 <- HPT.1 %>% filter(pregnant!="pregnant/PP")

## Summarize the data by sub-groups/ variables of interest
HPT.1 <- HPT.1%>%
  group_by(Author, Setting, Province,Study_Period, `Diagnostic criteria`, ROB, ART_status)%>%
  summarise(N=sum(N), size=sum(size))
View(HPT.1) # view the dataset

##### remove articles that did not reported ART status and those where ART status could not be deduced
HPT.ART <- HPT.1 %>% filter(ART_status!="unreported" & ART_status!="initiated and uninitiated")


##### meta analysis for hypertension:  studies sub grouped by ART status (intiated vs not initiated)
pool.HPT.ART <- metaprop(event = HPT.ART$N,n=HPT.ART$size,
                          fixed = F, random = T, method = "Inverse", method.tau = "DL",
                          sm = "PLOGIT", hakn = T, studlab = HPT.ART$Author,
                          overall = T, overall.hetstat = T,
                          method.bias = "Egger",backtransf = T,
                          text.random = "RE Prevalence",
                          subgroup = HPT.ART$ART_status,
                          test.subgroup = T, pscale = 100, )
summary(pool.HPT.ART)


###forest plot for HPT: studies sub grouped by ART status (intiated vs not initiated)

forest_ARTstatus_HPT <- forest.meta(pool.HPT.ART,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI]"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,100), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

#save forest plot
png(file = "ART_forest-HPT.png", width = 2800, height = 2800, res = 300)
forest_ARTstatus_HPT
dev.off()


##### meta analysis for hypertension subgroup diagnostics
pool.HPT.diag <- metaprop(event = HPT.1$N,n=HPT.1$size,
                         fixed = F, random = T, method = "Inverse", method.tau = "DL",
                         sm = "PLOGIT", hakn = T, studlab = HPT.1$Author,
                         overall = T, overall.hetstat = T,
                         method.bias = "Egger",backtransf = T,
                         text.random = "RE Prevalence",
                         subgroup = HPT.1$`Diagnostic criteria`,
                         test.subgroup = T, pscale = 100, )
summary(pool.HPT.diag)


###forest plot for HPT subgroup diagnostics

forest_diagnostic_HPT <- forest.meta(pool.HPT.diag,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,100), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

#save forest plot
png(file = "Diagnostics_forest-HPT.png", width = 2800, height = 2800, res = 300)
forest_diagnostic_HPT
dev.off()



##### meta analysis for hypertension: studies  subgrouped by setting (urban, rural)
## filter out by studies the were both in urban and rural settings
HPT.setting <- HPT.1%>%filter(Setting!="Urban_Rural")

pool.HPT.3 <- metaprop(event = HPT.setting$N,n=HPT.setting$size,
                      fixed = F, random = T, method = "Inverse", method.tau = "DL",
                      sm = "PLOGIT", hakn = T, studlab = HPT.setting$Author,
                      overall = T, overall.hetstat = T,
                      method.bias = "Egger",backtransf = T,
                      text.random = "RE Prevalence",
                      subgroup = HPT.setting$Setting,
                      test.subgroup = T, pscale = 100, )
summary(pool.HPT.3)

###forest plot for HPT subgroup setting

forest_setting_HPT <- forest.meta(pool.HPT.3,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,100), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
# save forest plot
png(file = "forest-HPT_2.png", width = 2800, height = 2800, res = 300)
forest_setting_HPT
dev.off()


##### meta analysis for hypertension subgroup STUDY PERIOD
pool.HPT.4 <- metaprop(event = HPT.1$N,n=HPT.1$size,
                       fixed = F, random = T, method = "Inverse", method.tau = "DL",
                       sm = "PLOGIT", hakn = T, studlab = HPT.1$Author,
                       overall = T, overall.hetstat = T,
                       method.bias = "Egger",backtransf = T,
                       text.random = "RE Prevalence",
                       subgroup = HPT.1$Study_Period,
                       test.subgroup = T, pscale = 100, )
summary(pool.HPT.4)

###forest plot for HPT subgroup STUDY PERIOD

forest_studyperiod_HPT <- forest.meta(pool.HPT.4,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,60), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

#save forest plot
png(file = "forest-HPT_period.png", width = 2800, height = 2700, res = 300)
forest_studyperiod_HPT
dev.off()


### ##### meta analysis for hypertension subgroup by sex

HPT_sex<- NCD.dat%>%filter(NCD=="HPT", !is.na(Sex))
View(HPT_sex)
HPT_sex <- HPT_sex%>%filter(pregnant!="pregnant/PP")

pool.HPT.sex <- metaprop(event = HPT_sex$N,n=HPT_sex$size,
                        fixed = F, random = T, method = "Inverse", method.tau = "DL",
                        sm = "PLOGIT", hakn = T, studlab = HPT_sex$Author,
                        overall = T, overall.hetstat = T,
                        method.bias = "Egger",backtransf = T,
                        text.random = "RE Prevalence",
                        subgroup = HPT_sex$Sex,
                        test.subgroup = T, pscale = 100, )
summary(pool.HPT.sex)

## forest plot
forest_sex_HPT <- forest.meta(pool.HPT.sex,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(16,"mm"),
            lwd = 2,colgap.left = unit(1,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(0,80), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")

# save forest plot
png(file = "forest-HPT_sex.png", width = 2800, height = 3300, res = 300)
forest_sex_HPT
dev.off()
```


## OBESITY
```{r}
##### subset out articles on OBESITY
obes.1 <- NCD.dat%>%filter(NCD=="obesity")

## remove studies on pregnant/postpartum participants

obes.1 <- obes.1%>%filter(pregnant!="pregnant/PP")

## Summarize the data by sub-groups/ variables of interest
obes.1 <- obes.1%>%
  group_by(Author, Setting, Province,Study_Period, ROB, ART_status)%>%
  summarise(N=sum(N), size=sum(size))
View(obes.1) # view the data


##### ##### remove articles that did not reported ART status and those where ART status could not be deduced
obes.ART <- obes.1 %>% filter(ART_status!="unreported" & ART_status!="initiated and uninitiated")

##### meta analysis for obesity subgroup ART status
pool.obes.ART<- metaprop(event = obes.ART$N,n=obes.ART$size,
                         fixed = F, random = T, method = "Inverse", method.tau = "DL",
                         sm = "PLOGIT", hakn = T, studlab = obes.ART$Author,
                         overall = T, overall.hetstat = T,
                         method.bias = "Egger",backtransf = T,
                         text.random = "RE Prevalence",
                         subgroup = obes.ART$ART_status,
                         test.subgroup = T, pscale = 100, )
summary(pool.obes.ART)


###forest plot for obesity subgroup ART

forest_ARTstatus_obsesity <-forest.meta(pool.obes.ART,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI]"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,60), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
#save forest plot
png(file = "ART_forest-obes.png", width = 2800, height = 2600, res = 300)
forest_ARTstatus_obsesity
dev.off()


### Remove studies conducted in both urban and rural settings
obes.setting <- obes.1%>%filter(Setting!="Urban_Rural")

##### meta analysis for obesity subgroup setting (urban, rural)
pool.obes.3 <- metaprop(event = obes.setting$N,n=obes.setting$size,
                       fixed = F, random = T, method = "Inverse", method.tau = "DL",
                       sm = "PLOGIT", hakn = T, studlab = obes.setting$Author,
                       overall = T, overall.hetstat = T,
                       method.bias = "Egger",backtransf = T,
                       text.random = "RE Prevalence",
                       subgroup = obes.setting$Setting,
                       test.subgroup = T, pscale = 100, )
summary(pool.obes.3)

###forest plot for obesity subgroup setting

forest_setting_obesity <- forest.meta(pool.obes.3,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,60), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
#save forest plot
png(file = "forest-obes_setting.png", width = 2800, height = 2600, res = 300)
forest_setting_obesity
dev.off()


##### meta analysis for obesity subgroup STUDY PERIOD
pool.obes.4 <- metaprop(event = obes.1$N,n=obes.1$size,
                        fixed = F, random = T, method = "Inverse", method.tau = "DL",
                        sm = "PLOGIT", hakn = T, studlab = obes.1$Author,
                        overall = T, overall.hetstat = T,
                        method.bias = "Egger",backtransf = T,
                        text.random = "RE Prevalence",
                        subgroup = obes.1$Study_Period,
                        test.subgroup = T, pscale = 100, )

summary(pool.obes.4)

###forest plot for obesity subgroup study period

forest_studyperiod_obesity <- forest.meta(pool.obes.4,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(13,"mm"),
            lwd = 2,colgap.left = unit(3,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(-10,60), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
# save forest plot
png(file = "forest-obes_period.png", width = 2800, height = 2600, res = 300)
forest_studyperiod_obesity
dev.off()


### meta analysis for obesity subgroup by sex

obes_sex<- NCD.dat%>%filter(NCD=="obesity", !is.na(Sex))

obes_sex <- obes_sex%>%filter(pregnant!="pregnant/PP") #remove pregnant people

## meta analysis by sex
pool.obes.sex <- metaprop(event = obes_sex$N,n=obes_sex$size,
                         fixed = F, random = T, method = "Inverse", method.tau = "DL",
                         sm = "PLOGIT", hakn = T, studlab = obes_sex$Author,
                         overall = T, overall.hetstat = T,
                         method.bias = "Egger",backtransf = T,
                         text.random = "RE Prevalence",
                         subgroup = obes_sex$Sex,
                         test.subgroup = T, pscale = 100, )
summary(pool.obes.sex)

## forest plot
forest_sex_obesity <- forest.meta(pool.obes.sex,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(16,"mm"),
            lwd = 2,colgap.left = unit(1,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(0,70), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
## save forest plot
png(file = "forest-obes_sex.png", width = 2800, height = 3300, res = 300)
forest_sex_obesity
dev.off()


##### Meta analysis comparing obesity prevalence in studies which included pregnant people and those the considered non-pregnant people

obes_preg<- NCD.dat%>%filter(NCD=="obesity",Sex=="Females", !is.na(pregnant))
View(obes_preg)

pool.obes.preg <- metaprop(event = obes_preg$N,n=obes_preg$size,
                          fixed = F, random = T, method = "Inverse", method.tau = "DL",
                          sm = "PLOGIT", hakn = T, studlab = obes_preg$Author,
                          overall = T, overall.hetstat = T,
                          method.bias = "Egger",backtransf = T,
                          text.random = "RE Prevalence",
                          subgroup = obes_preg$pregnant,
                          test.subgroup = T, pscale = 100, )
summary(pool.obes.preg)


# forest plot comparing obesity prevalence in studies which included pregnant people and those the considered unpregnant people
forest_pregnant_obesity <- forest.meta(pool.obes.preg,sortvar = event, studlab =T,
            layout = "RevMan5",backtransf =T, digits = 1,
            fs.random = 10, fs.heading = 10, fs.study = 10,
            leftcols =c("studlab","n","effect","ci"), 
            leftlabs = c("Author,Year","Sample","Prevalence %","[95% CI"),
            fs.test.overall = 10, fs.hetstat = 10, fs.test.effect.subgroup = 10,
            squaresize = 0.5,rightcols = F, colgap.forest.left = unit(16,"mm"),
            lwd = 2,colgap.left = unit(1,"mm"),
            smlab = "Prevalence %", just = "center", pscale = 100,
            xlim =c(0,70), col.study = "black",
            col.square = "black",col.square.lines = "black",
            col.diamond = "red",col.diamond.lines = "red",col.diamond.random = "red",
            col.label.left = "black",ff.random.labels = "bold",
            ff.hetstat = "bold",ff.test.subgroup = "bold",ff.study.labels = "plain",
            col.by = "black",
            text.random.w = "Subgroup Prevalence", text.random = "Overall Prevalence")
## save forest plot
png(file = "forest-obes_preg.png", width = 2800, height = 3300, res = 300)
forest_pregnant_obesity
dev.off()

```



